Density dependence

### 2020-09-16 R/density dependence.Rmd
checked convergence issue and cleaned up code.  Added 1-ha fragments and habitat as a predictor

```{r data, echo=TRUE}
ha <- read_rds(here("analysis", "data", "derived_data", "ha_density.rds"))
ha
```
Actually interested in density in the previous year

```{r}
ha <-
  ha %>% 
  group_by(ha_id_number) %>% 
  arrange(ha_id_number, year) %>% 
  mutate(across(starts_with("density"), lag)) %>% 
  ungroup() %>% 
  #scale year
  mutate(year_post = year - min(year))
```

@@

```{r}
ha_cf <- 
  ha %>% 
  filter(habitat == "CF") %>% 
  mutate(log_ht = log(ht),
         log_ht_prev = log(ht_prev))
```

@@

@@

```{r}
mpois <- glm(shts ~ log(shts_prev) * density_0_10,
             family = poisson(link = "log"),
             data = ha_cf)
lm <- lm(shts ~ log(shts_prev) * density_0_10,
         data = ha_cf)
mnorm <- glm(shts ~ log(shts_prev) * density_0_10,
             family = gaussian(link = "log"),
             data = ha_cf)
mpois2 <- glmer(shts ~ log(shts_prev) * density_0_10 + (1|ha_id_number) + (1|year_post) + (1|plot),
                family = poisson(link = "log"),
                data = ha_cf)

mnorm2 <- glmer(shts ~ log(shts_prev) * density_0_10 + (1|ha_id_number) + (1|year_post) + (1|plot),
                family = gaussian(link = "log"),
                data = ha_cf)

AICtab(mpois, mnorm, lm, mpois2, mnorm2)
```

Lognormal distribution fits best, even after attempting to account of overdispersion with random effects.  I guess despite being discrete, this isn't really "count" data *per se*.

Model doesn't converge though.

```{r}
mnorm3 <- glmer(shts ~ log(shts_prev) * density_0_10 + (1|ha_id_number) + (1|year_post),
               family = gaussian(link = "log"),
               data = ha_cf)

AICtab(mnorm2, mnorm3)
```

Can't remove plot ID.

```{r}
mnorm4 <- glmer(shts ~ log(shts_prev) * density_0_10 + (1|year_post) + (1|plot),
               family = gaussian(link = "log"),
               data = ha_cf)
AICtab(mnorm2, mnorm4)
```

@@

```{r}
mnorm5 <- glmer(shts ~ log(shts_prev) * density_0_10 + (1|ha_id_number) + (1|plot),
               family = gaussian(link = "log"),
               data = ha_cf)
AICtab(mnorm2, mnorm3, mnorm4, mnorm5)
```

Can't remove year as random effect.

Inspect residuals:

```{r}
ha_aug <- augment(mnorm2)
h <-
  ggplot(ha_aug, aes(.resid)) +
  geom_histogram()

qq <-
  ggplot(ha_aug, aes(sample = .resid)) +
  geom_qq() + geom_qq_line()

resid<- 
  ggplot(ha_aug, aes(x = .fitted, y = .resid)) + geom_point() 
h/(qq|resid)
```
I don't love this, but it is a better fit than a poisson distribution.

@@

- ~~Density should be total shoot number~~
- ~~Should actually be density in previous year~~
- ~~Shoots should be poisson distribution, not log-transformed?~~
- incorporate 1-ha fragments and use habitat type as a predictor.
- Implement FLM approach to density dependence.

>>>>

```{r data, echo=TRUE}
ha <- read_rds(here("analysis", "data", "derived_data", "ha_density.rds"))
ha
```

```{r}
ha <-
  ha %>% 
  group_by(ha_id_number) %>% 
  arrange(ha_id_number, year) %>% 
  ungroup() %>% 
  #scale year
  mutate(year_post = year - min(year))
```

@@

```{r}
ha_cf <- 
  ha %>% 
  filter(habitat == "CF") %>% 
  mutate(log_shts = log(shts),
         log_shts_next = log(shts_next))
```

Or filter to just 1-ha and cf

```{r}
ha_sub <-
  ha %>% 
  filter(habitat %in% c("CF", "1-ha")) %>% 
  mutate(log_shts = log(shts),
         log_shts_next = log(shts_next),
         habitat = as.factor(habitat))
```

@@

# General model structure

For each vital rate I'll include the number of shoots and habitat as fixed effects and plant ID and year as random effects.  I'm not including plot since there are few plots in each habitat type, and preliminary modeling shows the variance for plot is the lowest of those three random effects.

```{r}
ha_sub %>% 
  group_by(habitat) %>% 
  summarize(length(unique(plot)))
```

@@

## Distribution

Number of shoots is discrete, but not *really* count data.  I'll try a poisson distribution and a lognormal distribution.

```{r}
mpois <- glmer(shts_next ~ log_shts + habitat + density_0_10 + density_10_20 + (1|ha_id_number) + (1|year_post),
                family = poisson(link = "log"),
                data = ha_sub)

check_overdispersion(mpois2) #actually underdispersed

mnorm <- glmer(shts_next ~ log_shts + habitat + density_0_10 + density_10_20 + (1|ha_id_number) + (1|year_post),
                family = gaussian(link = "log"),
                data = ha_sub)
```

But how bad is this convergence failure?

Is the max|grad| calculation correct?

```{r}
derivs1 <- mnorm@optinfo$derivs
sc_grad1 <- with(derivs1, solve(Hessian, gradient))
max(abs(sc_grad1))
```
No, it's actually a lot smaller than `tol`, so nothing to worry about

```{r}
AICtab(mpois, mnorm)
compare_performance(mpois, mnorm, rank = TRUE)
```

Lognormal distribution fits best according to AIC, although poisson has lower RMSE.  I guess despite being discrete, this isn't really "count" data *per se*.  The lognormal model doesn't have problems with singularity that the poisson model has.  Both have problems with heteroscedasticity.  I'll go with log-normal.

```{r}
m_growth <- mnorm
```

```{r}
x<-check_heteroscedasticity(m_growth)
plot(x)
```

## Random effects

```{r}
m_growth2 <- glmer(shts_next ~ log_shts + habitat + density_0_10 + density_10_20 + (1|year_post),
               family = gaussian(link = "log"),
               data = ha_sub)
AICtab(m_growth, m_growth2)
```

@@

```{r}
m_growth3 <- glmer(shts_next ~ log_shts + habitat + density_0_10 + density_10_20 + (1|ha_id_number),
               family = gaussian(link = "log"),
               data = ha_sub)
AICtab(m_growth, m_growth2, m_growth3)
```

## Does density matter beyond immediate surroundings?

```{r}
Anova(m_growth)
```

Odd, density further out is significant, but density close is not.

```{r}
fixef(m_growth) %>% exp()
```

Both weakly positive density dependence

# Survival

Pre-reproductive census, so plants live/die in same year as `shts` and `ht`

```{r}
m_surv <- glmer(surv ~ log_shts + habitat + density_0_10 + density_10_20 + (1|ha_id_number) + (1|year), 
                family = binomial, data = ha_sub)
```

## Random effects

```{r}
m_surv2 <- glmer(surv ~ log_shts + habitat + density_0_10 + density_10_20 + (1|year), 
                family = binomial, data = ha_sub)
AICtab(m_surv, m_surv2)
```

Close to equivalent, so remove plant ID.  This makes sense, as plants only die once!

```{r}
m_surv3 <- glm(surv ~ log_shts + habitat + density_0_10 + density_10_20, 
                family = binomial, data = ha_sub)
AICtab(m_surv2, m_surv3)
```
Can't remove year.

## Density dependence?

```{r}
Anova(m_surv2)
fixef(m_surv2) %>% plogis()
```

Marginal, and only immediate grid cell.  Density has positive affect on survival.  So, probably just good patches.

# Flowering probability

```{r}
ggplot(ha_sub, aes(infl)) + geom_histogram()
```
Make a binary column

```{r}
ha_sub <- 
  ha_sub %>% 
  mutate(flwr = as.integer(infl>0), .before = infl)

#the way suvival is currently coded, a plant can flower and then not survive in the same year, so no need to filter out dead plants.
ha_sub %>% filter(surv == 0 & flwr == 1)
```

```{r}
m_flwr <- glmer(flwr ~ log_shts + habitat + density_0_10 + density_10_20 + (1|ha_id_number) + (1|year),
                family = binomial, data = ha_sub)
```

## Random effects

```{r}
m_flwr2 <- glmer(flwr ~ log_shts + habitat + density_0_10 + density_10_20 + (1|year),
                 family = binomial, data = ha_sub)
AICtab(m_flwr, m_flwr2)
```
Can't remove plant ID.

```{r}
m_flwr3 <- glmer(flwr ~ log_shts + habitat + density_0_10 + density_10_20 + (1|ha_id_number),
                 family = binomial, data = ha_sub)
AICtab(m_flwr, m_flwr2, m_flwr3)
```

Can't remove year

## Density dependence?

```{r}
Anova(m_flwr)
fixef(m_flwr) %>% plogis()
```

Ok, there's an effect, but oddly only past 20m.  Again, positive density dependence.

@@

- Check number of inflorescencesa given flowered?
- ~~check that survival is coded right so that it is a function of `ht_prev`~~
- try negative binomial distribution for shoot number
- ~~Density should be total shoot number~~
- ~~Should actually be density in previous year~~
- ~~Shoots should be poisson distribution, not log-transformed?~~
- ~~incorporate 1-ha fragments and use habitat type as a predictor.~~
- Implement FLM approach to density dependence.
